---
title: "Introduction to Logistic LASSO Regression"
author: "Johnson Phosavanh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## What is Logisitc Regression?
Logistic regression is binary classifier that outputs a probability which can be mapped to a class, where the outcome is encoded as 0s and 1s.

Logistic regression comes under the family of generalised linear models (GLMs) by extending a linear model to model a response variable via an invertible link function and a probability distribution. In the case of logistic regression, we have the following

- a linear predictor
$$g(\mathbf{x}_i)=\beta_0+\beta_1x_{i1}+\beta_2x_{i2}+\dots+\beta_px_{ip},$$
- a link function (the logistic function)
$$\pi_i=\frac{1}{1+\exp(-g(\mathbf{x}_i))}=\frac{\exp(g(\mathbf{x}_i))}{1+\exp(g(\mathbf{x}_i))},$$
- a probababilistic distribution
$$Y_i|X_i=\mathbf{x}_i\sim \text{Bernoulli}(\pi_i).$$

The fitted value for an observed $\mathbf{x}_0$ is
$$\hat{Y}=\frac{1}{1+\exp\left[-\left(\hat{\beta}_0+\hat{\beta}_1x_{01}+\hat{\beta}_2x_{02}+\dots+\hat{\beta}_px_{0p}\right)\right]}.$$

### How do we find $\hat{\boldsymbol{\beta}}$?


### How do I fit a logistic regression model in R?

```{r, eval = F}
glm(y ~ x, data = dat, family = binomial)
```


### Issues with logisitc regression

## LASSO Logistic Regression























